Evaluation

### 2019-02-27 evaluation.Rmd
testing the processing in of forecasts for evaluation

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
'%>%' <- magrittr::'%>%'
source("tools/forecast_tools.R")

model_metadata = yaml::yaml.load_file("data/metadata.yaml")
pred_interval = model_metadata$pred_interval
models_of_interest = c('AutoArima','Ensemble','ESSS','nbGARCH','pevGARCH')

species_of_interest = c('BA','DM','DO','PP','OT','NA','total')
species_names = read.csv('PortalData/Rodents/Portal_rodent_species.csv',
                         stringsAsFactors=FALSE, na.strings = '') %>%
  dplyr::select(species = speciescode, full_species_name = scientificname) %>%
  dplyr::add_row(species='total', full_species_name='Total Rodents')

new_moons = read.csv('PortalData/Rodents/moon_dates.csv') %>%
            dplyr::mutate(censusdate = as.Date(censusdate))

# Setup rodent observations
species_abundance = portalr::abundance(clean=FALSE, shape='flat', level='treatment', 
                                       plots="longterm", time='newmoon') %>%
  dplyr::filter(treatment=="control") %>%
  dplyr::rename(actual = abundance, level = treatment) %>%
  dplyr::mutate(level='Controls',currency='abundance') %>%
  dplyr::filter(species %in% species_of_interest)

total_abundance = portalr::abundance(clean=FALSE, shape='flat', level='treatment', 
                                     plots="longterm", time='newmoon') %>%
  dplyr::filter(treatment=="control") %>%
  dplyr::group_by(newmoonnumber) %>%
  dplyr::summarise(actual=sum(abundance)) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(level='Controls',currency='abundance',species='total')

observation_data = species_abundance %>%
  dplyr::bind_rows(total_abundance) %>% 
  dplyr::left_join(new_moons, by='newmoonnumber')

```

@@

This graph shows the forecast error of some recent forecasts for the most recent sampling dates. This is shown for all model types in the columns and several common rodent species (and total rodents) in the rows. These evaluations are for the Control plots where no manipulation is done. 
```{r recent_eval, echo=FALSE, message=FALSE, warning=FALSE,, fig.width=15, fig.height=15}
###############################################
#Get all the recent forecasts
forecast_data = compile_forecasts(use_hindcasts = FALSE) %>%
  dplyr::filter(model %in% models_of_interest)

# Round so that large decimal place differences do not cause
# similar forecasts to be seen as distinct. 
forecast_data$estimate = round(forecast_data$estimate, 3)
forecast_data$LowerPI = round(forecast_data$LowerPI, 3)
forecast_data$UpperPI = round(forecast_data$UpperPI, 3)

# Keep 1 forecast per unique initial_newmoon. When there are multiple forecasts
# using the same initial_newmoon, use the first one.
forecast_dates_to_keep = forecast_data %>%
  dplyr::select(date, initial_newmoon) %>%
  dplyr::distinct() %>%
  dplyr::group_by(initial_newmoon) %>%
  dplyr::filter(date == min(date)) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(keep='yes')

forecast_data = forecast_data %>%
  dplyr::left_join(forecast_dates_to_keep, by=c('date','initial_newmoon')) %>%
  dplyr::filter(keep=='yes') %>%
  dplyr::select(-keep)

##################################
#Combine forecasts and oservations
forecast_errors = forecast_data %>%
  dplyr::left_join(observation_data, by=c('newmoonnumber','level','species','currency')) %>% 
  dplyr::mutate(rmse = sqrt((estimate - actual)^2))

# Sometimes there are model runs on the same day and with all the same info,
# this gets rid of those
forecast_errors = forecast_errors %>%
  dplyr::distinct()

# Drop any entries that don't have an observation
# (ie. a rainy sample period)
forecast_errors = forecast_errors %>%
  dplyr::filter(!is.na(actual))

# Filter to common species and apply the full name
forecast_errors = forecast_errors %>%
  dplyr::filter(species %in% species_of_interest) %>%
  dplyr::left_join(species_names, by='species') %>%
  dplyr::select(-species) %>%
  dplyr::rename(species=full_species_name)

# Add a newline to the longer species names
forecast_errors$species = with(forecast_errors, ifelse(nchar(species)>15, stringr::str_replace(species,' ','\n'), species))

#############################################################
# Displaying the data from many forecasts in the past can create a mess of a plot. see #218.
# So here three forecasts are chosen from 1, 6, and 12 newmoons into the past. 
# Sometimes forecasts data is missing for a particular initial_newmoon, the recursive function 
# update_initial_newmoon attempts to find another one from the same time pluts/minus 2 month period.
#  This ensures 3 forecasts are always shown.
lead_times_to_show = c(1,6,12)
most_recent_observed_newmoon = max(new_moons$newmoonnumber)
initial_newmoons_to_keep = most_recent_observed_newmoon - lead_times_to_show

initial_newmoons_to_keep = sapply(initial_newmoons_to_keep, update_initial_newmoon)

forecast_errors = forecast_errors %>%
  dplyr::filter(initial_newmoon %in% initial_newmoons_to_keep)

#####################################
ggplot2::ggplot(forecast_errors, ggplot2::aes(x=censusdate, y=rmse, group=as.character(date), color=as.character(date))) +
  ggplot2::geom_point(size=4)+
  ggplot2::geom_line(size=1.5) +
  ggplot2::scale_color_brewer(palette='Dark2') +
  ggplot2::scale_x_date(breaks = unique(forecast_errors$censusdate)) +
  ggplot2::facet_grid(species~model, scales = 'free_y') + 
  ggplot2::theme_bw() +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle=45, hjust=0.99, debug = FALSE),
        axis.text = ggplot2::element_text(size=12),
        axis.title = ggplot2::element_text(size=20),
        strip.text = ggplot2::element_text(size=12),
        legend.text = ggplot2::element_text(size=15),
        legend.title = ggplot2::element_text(size=18)) +
  ggplot2::labs(x='Sample Date',y='RMSE', color='Forecast Issue\nDate')

```

@@

**RMSE**: Root mean square error, this is a metric used to evaluate the point estimate of a forecast.  
**Coverage**: This is the percentage of observations which fell within the 90% confidence interval of a forecast. Ideally this would be equal to 0.90. If it is higher than 0.90 the forecasts intervals are too wide, if it's lower then the forecast intervals are too narrow.

```{r hindcast_eval, echo=FALSE, message=FALSE, warning=FALSE,, fig.width=9, fig.height=15}

#Get the all the forecasts made during observation period
forecast_data = compile_forecasts(use_hindcasts = TRUE)
forecast_errors = calculate_forecast_error(observation_data, forecast_data, error_metric = 'RMSE') %>%
  dplyr::filter(error_value < 200) %>% #Drop RMSE greater than this because it throws off all the graphs
  dplyr::bind_rows(calculate_forecast_error(observation_data, forecast_data, error_metric = 'coverage'))

forecast_errors = forecast_errors %>%
  dplyr::left_join(species_names, by='species')

ggplot2::ggplot(forecast_errors, ggplot2::aes(x=lead_time, y=error_value, group=model, color=model)) +
  ggplot2::geom_point()+
  ggplot2::geom_line() +
  ggplot2::geom_hline(yintercept = 0.9) +
  ggplot2::labs(x='Lead Time (New Moons)') +
  ggplot2::facet_wrap(full_species_name~error_metric, scales = 'free_y', ncol=2) + 
  ggplot2::theme_bw() +
  ggplot2::labs(y = "Error Value", colour = "Model")

```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
  species <- rodent_spp(set = "evalplot")
  fcasts <- read_casts() %>%
            select_casts(species = species, level = "Controls") %>%
            append_observed_to_cast()
  fcast_errors <- measure_cast_error(fcasts)
  hcasts <- read_casts(cast_type = "hindcasts") %>%
            select_casts(species = species, level = "Controls") %>%
            append_observed_to_cast()
  hcast_errors <- measure_cast_error(hcasts)
```

## Metrics
**Error**: Error is the raw difference between a predicted value ($\hat{y}_{i}$) and observed value ($y_{i}$): 
**RMSE**: Root Mean Square Error (RMSE) is a metric used to evaluate the point estimate of a forecast.  
**Coverage**: Coverage is used to evaluate the uncertainty in a forecast and is the percentage of observations which fell within the confidence interval of the prediction. Because we use a 90% CI, coverage would ideally be equal to 0.90. If coverage is higher than 0.90, the forecast's intervals are too wide; if it's lower than 0.90, the forecast intervals are too narrow.

@@

@@
